---
title: "mamascubankitchen"
author: "Krystin sinclair"
date: "March 24, 2019"
output: html_document
---

Load in the two spreadsheets
1. Transactionial database
2. Supplemental data on event type and meal type
```{r loaddata}
transactions <- read.csv(file="mamascubankitchen_anonymized.csv", header=TRUE, sep=",")
library(readxl)
supplemental <- read.csv(file="mamascubankitchen_additionaldata.csv", header=TRUE, sep=",")
groupeditems <- read_excel("mamascubankitchen_groupedmenuitems.xlsx")
```
create new table that is grouped by date and has features of date, month, day of week) mealtype, location, event type, if a discount was offered, quantity sold of each menu item, total items sold, and sum of sales
```{r table}
#first step is to group the menu items
transaction_groupeditems <- merge(transactions, groupeditems, all.x = TRUE)
```
Lets deal with the negative values
```{r absolute}
transaction_groupeditems$Qty <- abs(transaction_groupeditems$Qty)
transaction_groupeditems$Net.Sales <- abs(transaction_groupeditems$Net.Sales)
transaction_groupeditems$Gross.Sales <- abs(transaction_groupeditems$Gross.Sales)
transaction_groupeditems$Tax <- abs(transaction_groupeditems$Tax)
```

```{r group}
library(magrittr) 
library(dplyr)
#we want to group by date and see total quanttty sold and total net sales
#create new dataframe of only date, quantity and net sales
#Group by date to get sum of qty and netsales for each date


totalsold <- transactions %>%
select(Date, Qty, Net.Sales) %>%
group_by(transactions$Date) %>%
summarise(Qty = sum(Qty), Net.Sales = sum(Net.Sales))

# drop observation with 0 items sold
totalsold<-subset(totalsold, Qty!=0)

#rename date column
names(totalsold)[1]<-"Date"
names(supplemental)[1] <-"Date"

#merge total sold with supplemental
totalsoldwithsupplementaldata <- merge(totalsold, supplemental, all.x = TRUE)

```

```{r category}
#create table with date as row index and named item as column headers with count in the cells
library(reshape)
library(reshape2)
library(rpivotTable)
#install.packages('rpivotTable')
items <- transaction_groupeditems[c("Date","nameditem", "Qty")]
# specify the row headings
#install.packages('pivottabler')
library(pivottabler)
pt <- PivotTable$new()
pt$addData(items)
pt$addColumnDataGroups("nameditem")
pt$addRowDataGroups("Date", addTotal=FALSE)  
pt$defineCalculation(calculationName="quantityofitems", summariseExpression="sum(Qty)")
pt$renderPivot()
items <- pt$asDataFrame()
```

```{r null}
#replace null values with 0 in pivot
#install.packages("imputeTS")
library("imputeTS")
items <- na.replace(items, 0)

#delete obs when there aree no sold items
# drop observation with 0 items sold
items<-subset(items, Total!=0)

#make date  a column instead of row names
library(tibble)
items<- items %>% rownames_to_column("Date")

```

```{r merge}
#merge the filled pivot with the totalsoldand supplemntal data
datatable <- merge(totalsoldwithsupplementaldata, items, all.x = TRUE)

#add in weather and join 
weather <- read_excel("weather.xlsx", sheet=2)

datatable$Date <- as.Date(datatable$Date, format = "%m/%d/%Y")
weather$Date <- as.Date(weather$Date, format = "%m/%d/%Y")

#merge
dataframe <- merge(datatable, weather, all.x = TRUE)


```

```{r date}
#get month and day of week from date
library(lubridate)
dataframe$month <- month(dataframe$Date)
dataframe$dayofweek <-  weekdays(as.POSIXct(dataframe$Date), abbreviate = T)
```

```{r dropexcescolumn}
dataframe <- subset(dataframe, select=-Total)

```

We want to see the categorical variables and treat them correeclty for analysis
```{r factors}
dataframe$dayofweek <- factor(dataframe$dayofweek)
dataframe$month <- factor(dataframe$month)
```


```{r proportions}
proportion <- dataframe

#columns we want proportions on 

names <- c('beverage','chicken', 'cuban sandwhich', 'dessert', 'daily special', 'empanadas', 'pork', 'sides')

for (n in names)
{
  proportion[, n] <- proportion[, n]/proportion$Qty

}

```

Lets look at the proportion of each item across dates
```{r proportionalitems}
for (n in names)
{
  print(n)
  print(summary(proportion[, n]))
}

```
```{r plotitems}
library(ggplot2)
bevplot <- ggplot(dataframe, aes(x=Date, y=Net.Sales, colour=beverage))  + geom_point(size=2) 
bevplot

chickplot <- ggplot(dataframe, aes(x=Date, y=Net.Sales, colour=chicken))  + geom_point(size=2) 
chickplot

cubplot <- ggplot(dataframe, aes(x=Date, y=Net.Sales, colour=`cuban sandwhich`))  + geom_point(size=2) 
cubplot

dayplot <- ggplot(dataframe, aes(x=Date, y=Net.Sales, colour=`daily special`))  + geom_point(size=2) 
dayplot

dessplot <- ggplot(dataframe, aes(x=Date, y=Net.Sales, colour=dessert))  + geom_point(size=2) 
dessplot

emplot <- ggplot(dataframe, aes(x=Date, y=Net.Sales, colour=empanadas))  + geom_point(size=2) 
emplot

porkplot <- ggplot(dataframe, aes(x=Date, y=Net.Sales, colour=pork))  + geom_point(size=2) 
porkplot

sidesplot <- ggplot(dataframe, aes(x=Date, y=Net.Sales, colour=sides))  + geom_point(size=2) 
sidesplot

qtyplot <- ggplot(dataframe, aes(x=Date, y=Net.Sales, colour=Qty))  + geom_point(size=2) 
qtyplot

```

LEts see if net sales and qty distrubiton looks like
```{r linear}

summary(dataframe$Net.Sales)
hist(dataframe$Net.Sales)

summary(dataframe$Qty)
hist(dataframe$Qty)


```

Create a Sales dataframe that has teh information we need to predict future sales

```{r cleanup}
#first remove columns that we only know after the event
Sales <- subset(dataframe, select=-Qty)

#we want to know which items were on the menu for each date

# we dont know how much of each item we will sell, but we do know if wew will sell the item
for (n in names)
{
  Sales[, n] [Sales[, n] != 0] <- 1

}

#create new column that sums the menu items to see how many items were available on that date

Sales$items <- Sales$beverage + Sales$chicken + Sales$`cuban sandwhich`+ Sales$`daily special`+ Sales$dessert+ Sales$empanadas+ Sales$pork + Sales$sides

#drop the columns for specific menu items
Sales$beverage <- Sales$chicken <- Sales$`cuban sandwhich` <- Sales$`daily special`<- Sales$dessert <- Sales$empanadas <- Sales$pork <- Sales$sides <- NULL

Sales$items <- factor(Sales$items)
```
 Let's explore our DAtaframe
```{r exploration}
str(Sales)
summary(Sales)
```



We want to view how these features relate to our variable of interest Net.Sales
```{r Sales}

#create a list of colnames
 
features <- colnames(Sales)


for (n in features)
{
  plot(Sales[, n], Sales$Net.Sales, main=n, xlab= n, ylab="Net Sales ($)")

}
```

```{r twofeatureplot}
#plot date and event type
library(ggplot2)

a <- ggplot(Sales, aes(x=Date, y=Net.Sales, colour=location)) + geom_point(size=2) 
a

b <- ggplot(Sales, aes(x=Date, y=Net.Sales, colour=mealtype, shape=event.type)) + geom_point(size=2) 
b

c<- ggplot(Sales, aes(x=Date, y=Net.Sales, colour=Temperature, shape=event.type))  + geom_point(size=2) 
c

d<- ggplot(Sales, aes(x=Date, y=Net.Sales, colour=Precipitation, shape=event.type))  + geom_point(size=2) 
d

e<- ggplot(Sales, aes(x=Date, y=Net.Sales, colour=items, shape=event.type))  + geom_point(size=2) 
e

f<- ggplot(Sales, aes(x=Date, y=Net.Sales, colour=items, shape=discount))  + geom_point(size=2) 
f

g <- ggplot(Sales, aes(x=Date, y=Net.Sales, colour=mealtype, shape=discount)) + geom_point(size=2) 
g

h<- ggplot(Sales, aes(x=Date, y=Net.Sales, colour=Temperature, shape=discount))  + geom_point(size=2) 
h

i<- ggplot(Sales, aes(x=Date, y=Net.Sales, colour=Precipitation, shape=discount))  + geom_point(size=2) 
i

j<- ggplot(Sales, aes(x=Date, y=Net.Sales, colour=event.type, shape=discount))  + geom_point(size=2) 
j

```
```{r 2x2}
#we want to create a 2x2 table and group by discount and event type as these features appear to be strongly realted to sales

a=Sales%>%
  group_by(discount,event.type)%>%
  mutate(avg_sales=mean(Net.Sales))%>%
  select(discount,event.type,avg_sales)%>%
  unique()
class(a)
a

yesevent=a[3,3]
yesstreet =a[2,3]
nostreet=a[1,3]
noevent =a[4,3]
as.numeric(yesevent)
as.numeric(yesstreet)
as.numeric(nostreet)
as.numeric(noevent)
print(yesevent, yesstreet, nostreet, noevent)
#find percenage difference

streetdiscountdiff <- (yesstreet - nostreet) / yesstreet
streetdiscountdiff

eventdiscountdiff <- (yesevent - noevent) / noevent
eventdiscountdiff
```
Now, we want to see if the categorical variables have influence over Net Sales

```{r groupby}
GroupedMealtype <- Sales %>% group_by(mealtype) %>% summarise(
  Net.Sales = mean(Net.Sales))

GroupedLocation <- Sales %>% group_by(location) %>% summarise(
  Net.Sales = mean(Net.Sales))

Groupedeventtype <- Sales %>% group_by(event.type) %>% summarise(
  Net.Sales = mean(Net.Sales))

Groupeddiscount <- Sales %>% group_by(discount) %>% summarise(
  Net.Sales = mean(Net.Sales))

Groupedmonth <- Sales %>% group_by(month) %>% summarise(
  Net.Sales = mean(Net.Sales))

Groupeddayofweek <- Sales %>% group_by(dayofweek) %>% summarise(
  Net.Sales = mean(Net.Sales))

Groupeditems <- Sales %>% group_by(items) %>% summarise(
  Net.Sales = mean(Net.Sales))

```

We will use the 5% significant level for all ANOVA examples

```{r ANOVA}
# One Way Anova (Completely Randomized Design) mealtype
mealtypefit <- aov(Net.Sales ~ mealtype, data=Sales)
plot(mealtypefit)
summary(mealtypefit)

# One Way Anova (Completely Randomized Design) Location
locationfit <- aov(Net.Sales ~ location, data=Sales)
plot(locationfit)
summary(locationfit)

# One Way Anova (Completely Randomized Design) month type
monthfit <- aov(Net.Sales ~ month, data=Sales)
plot(monthfit)
summary(monthfit)

# One Way Anova (Completely Randomized Design) dayofweek
dayofweekfit <- aov(Net.Sales ~ dayofweek, data=Sales)
plot(dayofweekfit)
summary(dayofweekfit)

#One Way Anova (Completely Randomized Design) items
itemsfit <- aov(Net.Sales ~ items, data=Sales)
plot(itemsfit)
summary(itemsfit)
```

Based on the p-value of these ANOVA tests, the significant features are dayofweek
month
location
items

Let's try pair wise t-test to see which gropuings differ
```{r pairwise}
#pairwise.t.test(Sales$Net.Sales, Sales$location)
pairwise.t.test(Sales$Net.Sales, Sales$month, p.adjust.method = "bonf")
pairwise.t.test(Sales$Net.Sales, Sales$dayofweek, p.adjust.method = "bonf")
#pairwise.t.test(Sales$Net.Sales,Sales$items, p.adjust.method = "bonf")
```
for location and items we wnat to know what are the unique alues and how many observatiosn ofr each
```{r countobs}
#location
#str(Sales$location)
Locations <- Sales%>%
  group_by(location)%>%
  summarise(sales_mean=mean(Net.Sales))
Locations[order(-Locations$sales_mean),] 
plot(Locations$location,Locations$sales_mean)

```

```{r itemsimportance}

table(Sales$items)
itemsnumber <- Sales%>%
  group_by(items)%>%
  summarise(sales_mean=mean(Net.Sales))
itemsnumber[order(-itemsnumber$sales_mean),] 
plot(itemsnumber$items,itemsnumber$sales_mean)

```
Based on pariwise ttest only month really impacts net sales. In fact August and October differ as well as October and NOvember. The rest are not statisically significant at the 5% significance level of a difference of means test. 

August and november have lower mean net sales than october. 


Next we want to look at the binary variables and see if those impact Net Sales
```{r ttest}
# independent 2-group t-test
# where y is numeric and x is a binary factor
discounttest <- t.test(Sales$Net.Sales~Sales$discount) 
eventtypetest <- t.test(Sales$Net.Sales~Sales$event.type)

discounttest

eventtypetest

```

Both of these tests are statistically significant at the 5% significance level. 

When a discount is offered those dates have a statistically signficiant higher mean in Net. Sales. 

When it is an event and not just on the street those dates havea  statisicaly significant higher mean in Net Sales. 

Now we want to see if the continuous variables affect Net Sales

```{r continuous}
#correlation
cor(Sales$Temperature, Sales$Net.Sales)
cor(Sales$Precipitation, Sales$Net.Sales)

```
Neither Temperature nor Precipitation has a strong linear relationship with Sale.

Now we will try some linera models to use as estimations of which features relate to netsales. 
IN each successive model features are removed due to large p values. 

```{r linearmodel}
model <- subset(Sales, select=c("Net.Sales", "mealtype", "event.type", "discount", "Temperature", "Precipitation", "month", "dayofweek", "items"))

allin <- lm(Net.Sales ~ ., data=model)
summary(allin)

```

```{r model2}
model2 <- subset(Sales, select=c("Net.Sales", "event.type",  "Temperature", "month", "dayofweek", "items"))

model_2 <- lm(Net.Sales ~ ., data=model2)
summary(model_2)

```

```{r model3}
model3 <- subset(Sales, select=c("Net.Sales", "event.type",  "Temperature", "month", "items"))

model_3 <- lm(Net.Sales ~ ., data=model3)
summary(model_3)

```

```{r model4}
model4 <- subset(Sales, select=c("Net.Sales", "event.type",  "Temperature", "items"))

model_4 <- lm(Net.Sales ~ ., data=model4)
summary(model_4)

```

We also want to know if it is worht attending the event or going to the street for each date of transaction. We are saying that $300 is a mimum for it to be worth it. 

```{r logistic}
worthy <- Sales



worthy$worth <- ifelse(worthy$Net.Sales > 300, 
c(1), c(0)) 

logmodel <- subset(worthy, select=c("mealtype", "discount", "Precipitation", "location", "worth", "event.type",  "Temperature", "items", "month", "dayofweek"))

log <- glm(worth~., data=logmodel)
summary(log)

logmodel2 <- subset(worthy, select=c("worth", "event.type",  "Temperature", "items"))

log2 <- glm(worth~., data=logmodel2)
summary(log2)

```
Let's see how the menu items and number of quantity sold realte and sales

```{r quantittysold}
dataframe2 <- dataframe

dataframe2$beverage <- ifelse(dataframe$beverage > 0, 
c(1), c(0))

dataframe2$chicken <- ifelse(dataframe$chicken > 0, 
c(1), c(0))

dataframe2$dessert <- ifelse(dataframe$dessert > 0, 
c(1), c(0))

dataframe2$empanadas <- ifelse(dataframe$empanadas > 0, 
c(1), c(0))

dataframe2$pork <- ifelse(dataframe$pork > 0, 
c(1), c(0))

dataframe2$sides <- ifelse(dataframe$sides > 0, 
c(1), c(0))

dataframe2$`cuban sandwhich` <- ifelse(dataframe$`cuban sandwhich` >0,
c(1), c(0))

qtymodel <- subset(dataframe2, select=c("Qty", "beverage", "chicken", "cuban sandwhich","daily special", "dessert", "empanadas", "pork", "sides"))
qtymodell <- lm(Qty~ ., data=qtymodel)
summary(qtymodell)

salesmodel <- subset(dataframe2, select=c("Net.Sales", "beverage", "chicken", "cuban sandwhich","daily special", "dessert", "empanadas", "pork", "sides"))
salesmodell <- lm(Net.Sales~., data=salesmodel)
summary(salesmodell)

```
